---
title: "lab5"
author: "Han Sun, Xinyi Zhu"
date: "2020.4.13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(maps)
library(plyr)
```

```{r}
accident <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
```

1. are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?

```{r}
table(accident$DAY_WEEK)

```

Sunday has more accidents than others. Sunday has 5873 accidents.

2. what time of the day do accidents happen (use variable HOUR)?

```{r}
table(accident$HOUR)
accident %>% 
  filter(HOUR != 99) %>%
  ggplot(aes(x = HOUR)) + geom_bar() + geom_text(stat = "count", aes(label = stat(count)),vjust=1, color="white", size=2)
```

The accidents happen from 14 to 23 o'clock in a day. The most accidents happen at 18 o'clock.

3. what is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?

```{r}
accident %>% 
  filter(DRUNK_DR >= 1) %>%
  count()
```

The number of accidents with at least one drunk driver is 8769.

4. In the person table, dentify drivers (PER_TYP == 1, see FARS manual) and subset on them.

```{r}
drivers<-subset(person,PER_TYP == 1)
```

5. Join accident and person table (work out which variable(s) to use)

```{r}
connection <- accident %>%
  full_join(person, by = c("STATE", "ST_CASE","COUNTY", "DAY", "MONTH", "HOUR", "MINUTE"))

```


6. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results.
```{r}
library(lubridate)
connection$SEX <- as.factor(connection$SEX)

connection %>% 
  filter(SEX ==1 | SEX ==2) %>%
  filter( HOUR <= 24) %>%
  ggplot(connection, mapping=aes(x=HOUR, fill=wday(DAY_WEEK, label =TRUE) )) + 
  geom_bar() + facet_wrap(~SEX) +labs(fill = "Day of Week")  +xlab("hours")+ggtitle("Accidents by week,hour,and sex")

```


7. Now plot a choropleth map of accidents on a county level

Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration??s (GSA) publication
Use readxl::read_xlsx to read in the GLCs
```{r}
glcs <- readxl::read_xlsx("/Users/jessie/Desktop/ds202/ds202_lab5/GLCs.xlsx")
counties <- map_data('county')
counties1 <- counties
acc1 <- accident
glcs$STATE <- glcs$`State Code`
glcs$COUNTY <- glcs$`County Code`
glcs <- glcs%>% select("STATE", "State Name", "COUNTY", "County Name")
accident <- accident %>% select("STATE", "COUNTY", "ST_CASE")

geo1 <- accident %>% join(glcs)
```
```{r}
geo1 <- geo1 %>% distinct()
geo1$region <- geo1$`State Name`
geo1$subregion <-geo1$`County Name`
geo1 <- geo1 %>% select(-c("STATE","COUNTY","State Name","County Name"))
geo1$region <- tolower(geo1$region)
geo1$subregion <- tolower(geo1$subregion)
glcs_acc <- geo1 %>% group_by(region, subregion) %>% dplyr::summarise(numAcc =n())
final <- counties1 %>% join(glcs_acc)
```

8. Visualize the number of accidents in each county using a choropleth map. To do this, you need to join the accident data, GLCs, and appropriate map data.
```{r}
ggplot(final, aes(x=long, y= lat, fill = numAcc)) + geom_polygon(aes(group = group)) +xlab("Longitude") +ylab("Latitude") + labs(fill="Number of Accidents")
```


9. Can you find seasonal trends in accident occurrences in different states? Is summer or winter more dangerous? Why?
```{r}
acc1 <- acc1 %>% filter(WEATHER <98)
Weather <- acc1 %>% group_by(STATE, WEATHER) %>% dplyr::summarise(n =n())

ggplot(Weather, aes(x=STATE, y =n, fill = WEATHER)) + geom_bar(stat = "identity") +
  xlim(c(0,60)) +ylim(c(0,3500)) +ylab('Number of Accidents')
```
From this bar char, I noticed the darker months (January and Frbruray most likely) account for a larger chunk of yearly accidents. As during the winter weather, there are snows on the road. That makes sense for me that number of accidents larger in these two monthes. Moreover, the lightly blue colored months(December and November) would also have the same reason for winter weather conditions.
