---
title: "lab5"
author: "Han Sun, Xinyi Zhu"
date: "2020.4.13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

```{r}
accident <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
```

1. are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?

```{r}
table(accident$DAY_WEEK)
max(count(accident,DAY_WEEK))
```

Sunday has more accidents than others. Sunday has 5873 accidents.

2. what time of the day do accidents happen (use variable HOUR)?

```{r}
table(accident$HOUR)
accident %>% 
  filter(HOUR != 99) %>%
  ggplot(aes(x = HOUR)) + geom_bar() + geom_text(stat = "count", aes(label = stat(count)),vjust=1, color="white", size=2)
```

The accidents happen from 14 to 23 o'clock in a day. The most accidents happen at 18 o'clock.

3. what is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?

```{r}
accident %>% 
  filter(DRUNK_DR >= 1) %>%
  count()
```

The number of accidents with at least one drunk driver is 8769.

4. In the person table, dentify drivers (PER_TYP == 1, see FARS manual) and subset on them.

```{r}
person <- person %>% 
  filter(PER_TYP == 1)
```

5. Join accident and person table (work out which variable(s) to use)

```{r}
connection <- accident %>%
  full_join(person, by = c("STATE", "ST_CASE","COUNTY", "DAY", "MONTH", "HOUR", "MINUTE"))
```


6. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results.

7. Now plot a choropleth map of accidents on a county level

Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration¡¯s (GSA) publication
Use readxl::read_xlsx to read in the GLCs

8. Visualize the number of accidents in each county using a choropleth map. To do this, you need to join the accident data, GLCs, and appropriate map data.

9. Can you find seasonal trends in accident occurrences in different states? Is summer or winter more dangerous? Why?